Evaluating Cloud-Optimized HDF5 for NASA's ICESat-2 Mission

### 2024-10-09 _tex/paper.tex
Built site for gh-pages

The most recent version of the Hierarchical Data Format is HDF5, a
common archival format for n-dimensional scientific data; it has been
utilized to store valuable information from astrophysics to earth
sciences and everything in between. As flexible and powerful as HDF5 can
be, it comes with big tradeoffs when it's accessed from remote storage
systems.

@@

The test files were originally cloud optimized by ``repacking'' them,
using a relatively new feature in the HDF5 C API called ``paged
aggregation''. Page aggregation does 2 things: fisrt, it collects
file-level metadata from datasets and stores it on dedicated metadata
blocks at the front of the file; second, it forces the library to write
both data and metadata using these fixed-size pages. Aggregation allows
client libraries to read file metadata with only a few requests using
the page size as a fixed request size, overriding the 1 request per
chunk behavior.

@@

As we can see in Figure~\ref{fig-2}, when we cloud optimize a file using
paged-aggregation there are some considerations and behavior that we had
to take into account. The first thing to observe is that page
aggregation will --as we mentioned-- consolidate the file-evel metadata
at the front of the file and will add information in the so-called
superblock\footnote{The HDF5 superblock is a crucial component of the
  HDF5 file format, acting as the starting point for accessing all data
  within the file. It stores important metadata such as the version of
  the file format, pointers to the root group, and addresses for
  locating different file components} The next thing to notice is that
page size us uses across the board for metadata and data as of October
2024 and version 1.14 of the HDF5 library, page size cannot dynamically
adjust to the total metadata size.

@@

\caption{\label{fig-4}shows that using paged aggregation alone is not a
complete solution. This behavior us caused by over-reads of data now
distributed in pages and the internals of HDF5 not knowing how to
optimize the requests. This means that if we cloud optimize alone and
use the same code, in some cases we'll make access to these files even
slower. A very important thing to notice here is that rechunking the
file, in this case using 10X bigger chunks results in a predictable 10X
improvement in access times without any cloud optimization involved.
Having less chunks generates less metadata and bigger requests, in
general is it recommended that chunk sizes should range between 1MB and
10MB{[}Add citation, S3 and HDF5{]} and if we have anough memory and
bandwith even bigger (Pangeo recommends up to 100MB chunks){[}Add
citation.{]}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  File page strategy: H5F\_FSPACE\_STRATEGY\_PAGE
\item
  File page size: 8000000 If repacking an existing file, h5repack
  contains the code to alter tese variables inside the file
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Avoid using unlimited dimensions when creating variables because the
  HDF5 API cannot support it inside bffered pages and representation of
  these variables is not supported by Kerchunk.
\end{enumerate}

>>>>

The most recent version of the Hierarchical Data Format is HDF5, a
common archival format for n-dimensional scientific data; it has been
utilized to store valuable information from astrophysics to earth
sciences and everything in between. As flexible and powerful as HDF5 can
be, it comes with big trade-offs when it's accessed from remote storage
systems.

@@

The test files were originally cloud optimized by ``repacking'' them,
using a relatively new feature in the HDF5 C API called ``paged
aggregation''. Page aggregation does 2 things: first, it collects
file-level metadata from datasets and stores it on dedicated metadata
blocks at the front of the file; second, it forces the library to write
both data and metadata using these fixed-size pages. Aggregation allows
client libraries to read file metadata with only a few requests using
the page size as a fixed request size, overriding the 1 request per
chunk behavior.

@@

As we can see in Figure~\ref{fig-2}, when we cloud optimize a file using
paged-aggregation there are some considerations and behavior that we had
to take into account. The first thing to observe is that page
aggregation will --as we mentioned-- consolidate the file-level metadata
at the front of the file and will add information in the so-called
superblock\footnote{The HDF5 superblock is a crucial component of the
  HDF5 file format, acting as the starting point for accessing all data
  within the file. It stores important metadata such as the version of
  the file format, pointers to the root group, and addresses for
  locating different file components} The next thing to notice is that
page size us uses across the board for metadata and data as of October
2024 and version 1.14 of the HDF5 library, page size cannot dynamically
adjust to the total metadata size.

@@

\caption{\label{fig-4}shows that using paged aggregation alone is not a
complete solution. This behavior us caused by over-reads of data now
distributed in pages and the internals of HDF5 not knowing how to
optimize the requests. This means that if we cloud optimize alone and
use the same code, in some cases we'll make access to these files even
slower. A very important thing to notice here is that rechunking the
file, in this case using 10X bigger chunks results in a predictable 10X
improvement in access times without any cloud optimization involved.
Having less chunks generates less metadata and bigger requests, in
general is it recommended that chunk sizes should range between 1MB and
10MB{[}Add citation, S3 and HDF5{]} and if we have enough memory and
bandwidth even bigger (Pangeo recommends up to 100MB chunks){[}Add
citation.{]}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  File page strategy: H5F\_FSPACE\_STRATEGY\_PAGE
\item
  File page size: 8000000 If repacking an existing file, h5repack
  contains the code to alter these variables inside the file
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Avoid using unlimited dimensions when creating variables because the
  HDF5 API cannot support it inside buffered pages and representation of
  these variables is not supported by Kerchunk.
\end{enumerate}
